---
title: "Ethnicity and Identity"
subtitle: "Data Management"
author: "Michael Foley"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    highlight: haddock
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)

library(tidyverse)
library(janitor)
library(survey)
library(scales)
library(gtsummary)
```

The [American Trends Panel Wave 77](https://www.pewresearch.org/social-trends/dataset/american-trends-panel-wave-77/) data set is available for download at the PEW Research web site. This data was collected from a survey conducted Oct. 13-19, 2020. Everyone who took part is a member of the Pew Research Center’s American Trends Panel (ATP), an online survey panel that is recruited through national, random sampling of residential addresses. The survey is weighted to be representative of the U.S. adult population by gender, race, ethnicity, partisan affiliation, education and other categories ([ATP methodology](https://www.pewresearch.org/methods/u-s-survey-research/american-trends-panel/)).

```{r warning=FALSE, message=FALSE, class.source = 'fold-show'}
dat_0 <- foreign::read.spss(
  "../data/ATP W77.sav", 
  to.data.frame = TRUE
)

svy_0 <- svydesign(ids = ~1, weights = ~ WEIGHT_W77, data = dat_0)
```

The data set consists of `r dim(dat_0)[2]` variables collected from `r dim(dat_0)[1] %>% comma(1)` participants. Only a fraction of the columns are of interest for this analysis.

I did not find a codebook associated with the descriptive fields, so I'm going to have to infer a bit. The questions and associated variable names are in the [questionanaire](file:///C:/Users/mpfol/AppData/Local/Temp/Temp1_W77_Oct20.zip/W77_Oct20/ATP%20W77%20questionnaire.pdf).

## Demographics

Most demographic variables are self-explanatory from the variable name or from the response options. A few need explanation.

* `XKIDS1_W77`: PARENT OF CHILD <18, REGARDLESS OF WHETHER CHILD IS IN HOUSEHOLD.
* `XPARENT_W77`: PARENT WITH CHILD <18 IN HOUSEHOLD

```{r}
svy_0 %>% gtsummary::tbl_svysummary(
  include = c(starts_with("X"), starts_with("F_"), E_COVID_W77, 
              E_COVIDPARTNER_W77, WFHCURR_W77, WFHPARTNERCURR_W77,
              WFHPOSTCVD_W77, CHILD_LT12_W77))
```

## Questions

You can read all of the questions and response options in the [questionnaire](file:///C:/Users/mpfol/AppData/Local/Temp/Temp1_W77_Oct20.zip/W77_Oct20/ATP%20W77%20questionnaire.pdf). I am particularly interested in questions related to how spouses share family responsibilities. In each query below, I filter on `E_COVID_W77 %in% c("Employed full-time", "Employed part-time")`.

### `FAMSURV1_W77`

All survey participants were asked,

> At this point in your life, which of the following would be best for you personally? 

```{r class.source = 'fold-show'}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "FAMSURV1_W77", percent = "col")
```

### `FAMSURV32`

> How do you feel about each of the following?
> 
> * `FAMSURV32_a_W77`: The way household chores are divided between you and your spouse or partner
> * `FAMSURV32_c_W77`: How well you and your spouse or partner communicate with each other
> * `FAMSURV32_d_W77`: How well your spouse or partner balances work and their personal life
> * `FAMSURV32_e_W77`: Your spouse or partner’s approach to parenting
  
```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = starts_with("FAMSURV32"), percent = "col")
```

### `FAMSURV35`

> Since the beginning of the coronavirus outbreak, have you personally experienced the following because you were balancing work and parenting responsibilities?
> 
> * `FAMSURV35_a_W77`: Needed to reduce your work hours
> * `FAMSURV35_b_W77`: Felt like you couldn’t give 100% at work
> * `FAMSURV35_c_W77`: Turned down a promotion 
> * `FAMSURV35_d_W77`: Turned down an important assignment at work

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = starts_with("FAMSURV35"), percent = "col")
```

### `FAMSURV36`

> Since the beginning of the coronavirus outbreak, have you personally experienced the 
following at work because you have children?
>
>* `FAMSURV36_a_W77`: Being passed over for a promotion
>* `FAMSURV36_b_W77`: Being passed over for an important assignment
>* `FAMSURV36_c_W77`: Being treated as if you weren’t committed to your work

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = starts_with("FAMSURV36"), percent = "col")
```

### `FAMSURV38`

> For each of the following, who would you say does more?
> 
> * `FAMSURV38_a_W77`: Household chores and responsibilities
> * `FAMSURV38_b_W77`: Managing household finances
> * `FAMSURV38_c_W77`: Managing your child(ren)’s schedule and activities
> * `FAMSURV38_d_W77`: Being an involved parent

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = starts_with("FAMSURV38"), percent = "col")
```

### `TIMEOFF`

> Some companies have given working parents additional time off to deal with disruptions caused by the coronavirus outbreak. Do you think this is…

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "TIMEOFF_W77", percent = "col")
```

### `TIMECHILD`

> Thinking about the time *you* spend with your CHILDREN these days, do you think you spend…

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "TIMECHILD_W77", percent = "col")
```

### `TIMESPOUSE`

> Thinking about the time you spend with your SPOUSE OR PARTNER these days, do you think you spend…

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "TIMESPOUSE_W77", percent = "col")
```

### `CHILDHOME`

> Do you have any child care responsibilities when you’re working from home?

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "CHILDHOME_W77", percent = "col")
```

### `CVDBALANCE`

> Compared with before the coronavirus outbreak, is it...

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "CVDBALANCE_W77", percent = "col")
```

### `CVCHILDCARE`

> How easy or difficult has it been for you to handle child care responsibilities during the coronavirus outbreak?

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "CVCHILDCARE_W77", percent = "col")
```

### `WLBALANCE1`

>  How difficult, if at all, would you say it is for you personally to balance the responsibilities of your job with the responsibilities of your family?

```{r}
svy_0 %>% 
  subset(E_COVID_W77 %in% c("Employed full-time", "Employed part-time")) %>%
  tbl_svysummary(by = "F_GENDER", include = "WLBALANCE1_W77", percent = "col")
```

## Exploration: Validation with Reports

Pew produced several reports from this survey. To make sure I understand the data correctly, this section reproduces some of their important results.

```{r}
freq_by_gender <- function(dat, freq_var) {
  dat %>% 
    group_by(F_GENDER, !!sym(freq_var)) %>%
    summarize(.groups = "drop", wt = sum(WEIGHT_W77), n = n()) %>%
    group_by(F_GENDER) %>%
    mutate(pct = wt / sum(wt)) %>%
    ungroup()
}
```

### Figure #1: `CVCHILDCARE`

In Pew's Jan 2021 publication, [A rising share of working parents in the U.S. say it’s been difficult to handle child care during the pandemic](https://www.pewresearch.org/fact-tank/2021/01/26/a-rising-share-of-working-parents-in-the-u-s-say-its-been-difficult-to-handle-child-care-during-the-pandemic/), Pew writes 

> Overall, about half of employed parents with children younger than 12 in the household (52%) say it has been difficult to handle child care responsibilities during the coronavirus outbreak... Both working mothers and fathers with children younger than 12 are more likely than they were earlier in the pandemic to say it’s been difficult to handle child care responsibilities. But as was the case in March, larger shares of mothers than fathers say this (57% vs. 47%).

```{r fig.height=3.5, fig.width=6.5, warning=FALSE}
dat_working_parent <- dat_0 %>% 
  filter(
    # excludes "retired" and "not employed"
    E_COVID_W77 %in% c("Employed full-time", "Employed part-time") & 
    # excludes "More than one and do not consider any of them my primary job"
    JOBNUM_W77 %in% c("One", "More than one, but consider one of them my primary job") &
    # parent or guardian of any children under the age of 12 who live in household
    CHILD_LT12_W77 == "Yes" 
  )

# CVCHILDCARE_W77: How easy or difficult has it been for you to handle child 
# care responsibilities during the coronavirus outbreak?
fig1_dat <- dat_working_parent %>% 
  freq_by_gender("CVCHILDCARE_W77") %>%
  # refusing to answer excluded from display, but not from calc, so filter 
  # _after_ calculation.
  filter(CVCHILDCARE_W77 != "Refused" & F_GENDER %in% c("A man", "A woman")) %>%
  mutate(CVCHILDCARE_W77 = fct_drop(CVCHILDCARE_W77))

fig1_total <- dat_working_parent %>% 
  mutate(F_GENDER = "(total)") %>%
  freq_by_gender("CVCHILDCARE_W77") %>%
  filter(CVCHILDCARE_W77 != "Refused") %>%
  mutate(CHILDHOME_W77 = fct_drop(CVCHILDCARE_W77))

fig1_na <- fig1_total %>% mutate(F_GENDER = "NA", pct = NA_real_)

palette_fig1 <- c("darkseagreen", "darkseagreen2", "lightgoldenrod", "lightgoldenrod3")
names(palette_fig1) <- levels(fig1_dat$CVCHILDCARE_W77)

bind_rows(fig1_dat, fig1_total, fig1_na) %>%
  mutate(F_GENDER = factor(F_GENDER, levels = c("(total)", "NA", "A man", "A woman"),
                           labels = c("All Working Parents with Children under 12", 
                                      " ", "Fathers", "Mothers")),
         CVCHILDCARE_W77 = fct_relevel(CVCHILDCARE_W77, "Somewhat difficult", after = 3)) %>% 
  ggplot(aes(x = if_else(str_detect(CVCHILDCARE_W77, "easy"), -pct, +pct), 
             y = fct_rev(F_GENDER),
             fill = CVCHILDCARE_W77)) +
  geom_col() +
  geom_text(aes(label = percent(pct, 1)), position = position_stack(vjust = .5)) +
  geom_hline(yintercept = " ", linetype = 2, color = "lightgray") +
  geom_vline(xintercept = 0, linetype = 1, color = "lightgray") +
  scale_fill_manual(values = palette_fig1[1:4]) +
  scale_y_discrete(labels = wrap_format(18)) +
  theme_light() +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "Mothers are more likely to say handling child care is difficult.",
       subtitle = str_wrap("Among employed parents with children under 12 at home, % saying it has been ____ to handle child care responsibilities.", 70),
       caption = "Note: Share of respondents who did not answer not shown.")
```

### Figure #2: `CHILDHOME`

Next in Pew's Jan 2021 publication, Pew writes 

> Among employed parents who are working remotely all or most of the time and have children younger than 18 at home, about two-thirds (65%) say they have at least some child care responsibilities when working from home, including 26% who say they have a lot.
>
> Mothers and fathers who are working from home all or most of the time are equally likely to say they have at least some child care responsibilities while working from home (66% vs. 65%), though moms are about twice as likely as dads to say they have a lot of these duties (36% vs. 16%).

```{r fig.height=4.0, fig.width=6.5, warning=FALSE}
dat_working_parent_18_wfh <- dat_0 %>% 
  filter(
    # excludes "retired" and "not employed"
    E_COVID_W77 %in% c("Employed full-time", "Employed part-time") & 
    # excludes "More than one and do not consider any of them my primary job"
    JOBNUM_W77 %in% c("One", "More than one, but consider one of them my primary job") &
    # parent or guardian of any children under the age of 18 who live in household
    XPARENT_W77 == "Yes" &
    # working from home
    WFHCURR_W77 %in% c("All of the time", "Most of the time")
  ) 

# CVCHILDCARE_W77: How easy or difficult has it been for you to handle child 
# care responsibilities during the coronavirus outbreak?
fig2_dat <- dat_working_parent_18_wfh %>% 
  freq_by_gender("CHILDHOME_W77") %>%
  # refusing to answer excluded from display, but not from calc, so filter 
  # _after_ calculation.
  filter(CHILDHOME_W77 != "Refused" & F_GENDER %in% c("A man", "A woman"))  %>%
  mutate(CHILDHOME_W77 = fct_drop(CHILDHOME_W77))

fig2_total <- dat_working_parent_18_wfh %>% 
  mutate(F_GENDER = "(total)") %>%
  freq_by_gender("CHILDHOME_W77") %>%
  filter(CHILDHOME_W77 != "Refused") %>%
  mutate(CHILDHOME_W77 = fct_drop(CHILDHOME_W77))

fig2_na <- fig2_total %>% mutate(F_GENDER = "NA", pct = NA_real_)
fig2_na2 <- fig2_total %>% mutate(F_GENDER = "NA2", pct = NA_real_)
# fig2_na2$CHILDHOME_W77 <- c("Total", "Not much or None", "Some or a lot", "Total")
# fig2_na2$pct <- c(-.5, -.3, +.3, +.8)

fig2_posneg <- bind_rows(fig2_dat, fig2_total, fig2_na, fig2_na2) %>%
  mutate(CHILDHOME_W77 = fct_collapse(CHILDHOME_W77, 
                                      `Some or a lot` = c("Some", "A lot"),
                                      `Not much or None` = c("Not much", "None at all"))) %>%
  group_by(F_GENDER, CHILDHOME_W77) %>%
  summarise(.groups = "drop", across(wt:pct, sum)) %>%
  mutate(F_GENDER = factor(F_GENDER, levels = c("NA2", "(total)", "NA", "A man", "A woman"),
                           labels = c("  ", "All Teleworking Parents", 
                                      " ", "Fathers", "Mothers")))

palette_fig2 <- c("darkseagreen", "darkseagreen2", "lightgoldenrod", "lightgoldenrod3")
names(palette_fig2) <- levels(fig2_dat$CHILDHOME_W77)

bind_rows(fig2_dat, fig2_total, fig2_na, fig2_na2) %>%
  mutate(F_GENDER = factor(F_GENDER, levels = c("NA2", "(total)", "NA", "A man", "A woman"),
                           labels = c("  ", "All Teleworking Parents", 
                                      " ", "Fathers", "Mothers")),
         CHILDHOME_W77 = fct_relevel(CHILDHOME_W77, "Some", after = 3)) %>% 
  ggplot(aes(x = if_else(CHILDHOME_W77 %in% c("Not much", "None at all"), -pct, +pct), 
             y = fct_rev(F_GENDER),
             fill = CHILDHOME_W77)) +
  geom_col() +
  geom_text(aes(label = percent(pct, 1)), position = position_stack(vjust = .5)) +
  geom_text(data = fig2_posneg, aes(x = if_else(CHILDHOME_W77 == "Not much or None", -.5, +.8),
                                    label = percent(pct, 1)), size = 3, fontface = "bold") +
  geom_hline(yintercept = " ", linetype = 2, color = "lightgray") +
  geom_vline(xintercept = 0, linetype = 1, color = "lightgray") +
  scale_fill_manual(values = palette_fig2[1:4]) +
  scale_y_discrete(labels = wrap_format(18)) +
  theme_light() +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(x = NULL, y = NULL, fill = NULL,
       title = str_wrap("Teleworking mothers twice as likely to say they have a lot of child care duties while working.", 70),
       subtitle = str_wrap("Among teleworking parents with children under 18 at home, % saying they have ____ child care responsibilities when working from home.", 70),
       caption = "Note: Share of respondents who did not answer not shown.") +
  annotate("text", x = -.5, y = "  ", label = "Tot", size = 3, fontface = "bold",
           hust = 0, color = "darkgoldenrod") +
  annotate("text", x = +.8, y = "  ", label = "Tot", size = 3, fontface = "bold", 
           hust = 0, color = "darkseagreen")
```

### Figure #4: `FAMSURV35`

Pew writes 

> When asked about some of the specific challenges they may have faced at work since the beginning of the coronavirus outbreak because they were balancing work and parenting responsibilities, roughly half of employed parents (49%) say they felt like they could not give 100% at work since the outbreak began and 30% say they needed to reduce their work hours. 

```{r fig.height=4.5, fig.width=6.5, warning=FALSE}
dat_working_parent_18 <- dat_0 %>% 
  filter(
    # excludes "retired" and "not employed"
    E_COVID_W77 %in% c("Employed full-time", "Employed part-time") & 
    # excludes "More than one and do not consider any of them my primary job"
    JOBNUM_W77 %in% c("One", "More than one, but consider one of them my primary job") &
    # parent or guardian of any children under the age of 18 who live in household
    XPARENT_W77 == "Yes"
  ) 

# FAMSURV35_a_W77`: Needed to reduce your work hours
fig4_dat_a <- dat_working_parent_18 %>% freq_by_gender("FAMSURV35_a_W77")
fig4_total_a <- dat_working_parent_18 %>% mutate(F_GENDER = "(total)") %>% freq_by_gender("FAMSURV35_a_W77")
# FAMSURV35_b_W77`: Felt like you couldn’t give 100% at work
fig4_dat_b <- dat_working_parent_18 %>% freq_by_gender("FAMSURV35_b_W77") 
fig4_total_b <- dat_working_parent_18 %>% mutate(F_GENDER = "(total)") %>% freq_by_gender("FAMSURV35_b_W77")
# FAMSURV35_c_W77`: Turned down a promotion 
fig4_dat_c <- dat_working_parent_18 %>% freq_by_gender("FAMSURV35_c_W77") 
fig4_total_c <- dat_working_parent_18 %>% mutate(F_GENDER = "(total)") %>% freq_by_gender("FAMSURV35_c_W77")
# FAMSURV35_d_W77`: Turned down an important assignment at work
fig4_dat_d <- dat_working_parent_18 %>% freq_by_gender("FAMSURV35_d_W77") 
fig4_total_d <- dat_working_parent_18 %>% mutate(F_GENDER = "(total)") %>% freq_by_gender("FAMSURV35_d_W77")

fig4_dat <- bind_rows(
  fig4_dat_a %>% filter(FAMSURV35_a_W77 == "Yes, have experienced this" & F_GENDER %in% c("A man", "A woman")) %>%
    mutate(item = "Needed to reduce their work hours") %>% select(F_GENDER, item, pct),
  fig4_dat_b %>% filter(FAMSURV35_b_W77 == "Yes, have experienced this" & F_GENDER %in% c("A man", "A woman")) %>%
    mutate(item = "Felt like they couldn’t give 100% at work") %>% select(F_GENDER, item, pct),
  fig4_dat_c %>% filter(FAMSURV35_c_W77 == "Yes, have experienced this" & F_GENDER %in% c("A man", "A woman")) %>%
    mutate(item = "Turned down a promotion") %>% select(F_GENDER, item, pct),
  fig4_dat_d %>% filter(FAMSURV35_d_W77 == "Yes, have experienced this" & F_GENDER %in% c("A man", "A woman")) %>%
    mutate(item = "Turned down an important assignment at work") %>% select(F_GENDER, item, pct),
) %>%
  mutate(F_GENDER = fct_drop(F_GENDER))

fig4_tot <- bind_rows(
  fig4_total_a %>% filter(FAMSURV35_a_W77 == "Yes, have experienced this") %>%
    mutate(item = "Needed to reduce their work hours") %>% select(F_GENDER, item, pct),
  fig4_total_b %>% filter(FAMSURV35_b_W77 == "Yes, have experienced this") %>%
    mutate(item = "Felt like they couldn’t give 100% at work") %>% select(F_GENDER, item, pct),
  fig4_total_c %>% filter(FAMSURV35_c_W77 == "Yes, have experienced this") %>%
    mutate(item = "Turned down a promotion") %>% select(F_GENDER, item, pct),
  fig4_total_d %>% filter(FAMSURV35_d_W77 == "Yes, have experienced this") %>%
    mutate(item = "Turned down an important assignment at work") %>% select(F_GENDER, item, pct),
)

palette_fig4_fill <- c("darkseagreen", "lightgoldenrod")
names(palette_fig4_fill) <- levels(fig4_dat$F_GENDER)
palette_fig4_color <- c("darkseagreen4", "goldenrod")
names(palette_fig4_color) <- levels(fig4_dat$F_GENDER)

fig4_dat %>%
  ggplot(aes(x = pct, y = item)) +
  geom_point(aes(color = F_GENDER, fill = F_GENDER), size = 5, shape = 21) +
  geom_text(aes(label = percent(pct, 1)), vjust = -1, size = 3) +
  scale_fill_manual(values = palette_fig4_fill) +
  scale_color_manual(values = palette_fig4_color) +
  scale_y_discrete(labels = wrap_format(15)) +
  theme_light() +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color = "gray80", size = 1, linetype = 1),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  guides(color = FALSE) +
  labs(x = NULL, y = NULL, fill = NULL, color = NULL,
       title = str_wrap("Working mothers more likely to say they are struggling at work during pandemic.", 65),
       subtitle = str_wrap("Among employed parents with children under 18 at home, % saying that since the beginning of the coronavirus outbreak, they have personally exeperienced the following.", 75),
       caption = "Note: Share of respondents who did not answer not shown.")
```

## Save Data

Save the refined data set to a file for subsequent steps.

```{r}
saveRDS(dat_0, "../data/1_data_mgmt.rds")
```
